---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tidyusoc

<!-- badges: start -->
[![R-CMD-check](https://github.com/wklimowicz/tidyusoc/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/wklimowicz/tidyusoc/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Installation

You can install the development version of tidyusoc from [GitHub](https://github.com/) with:

```{r eval = FALSE}
# install.packages("devtools")
devtools::install_github("wklimowicz/tidyusoc")
```

## Setup

After downloading the Understanding Society Data from the UK Data Service website, run:

```{r example, eval = FALSE}
library(tidyusoc)

# Convert to Rds to save space and time
usoc_convert(usoc_directory = "spss/spss25",
             new_directory = "rds",
             filter_files = "indresp")

# Compile the INDRESP files in the rds folder
usoc_compile(directory = "rds")

```

A small set of core variables is included by default:

```{r echo = FALSE}
tibble::tribble(~Variable, ~Description,
                "pidp", "Personal ID",
                "age", "Age",
                "race", "Race/Ethnicity",
                "sex", "Sex",
                "country", "Country",
                "gor_dv", "Government Office Region",
                "qfhigh_dv", "Highest Qualification (Detailed, UKHLS Only)",
                "hiqual_dv", "Highest Qualification",
                "jbstat", "Job Status",
                "jbsoc00_cc", "Occupation",
                "fimnlabgrs_dv", "Labour Income" 
                ) |> kableExtra::kbl(format = "pipe")
```

To add more variables, use the `extra_mappings` function. For variables that change over time use `pick`. For example, number of A-Level's is:

* `qnqfedj` in waves BH1 to BH18
* `alevel` in waves 9 to 11

The variable BMI (`bmi_dv`) doesn't change over time, so it's passed as a string.

```{r eval = FALSE}

custom_mappings <- function(cols) {

    alevels <- pick(c("nqfedj", "alevel"), cols)

    custom_variables <- tibble::tribble(
        ~usoc_name, ~new_name, ~type,
        alevels, "alevels", "factor",
        "bmi_dv", "bmi_dv", "numeric"
    )

    return(custom_variables)

}

usoc <- usoc_compile("rds", extra_mappings = custom_mappings)
```
A good place to start looking for variables is the [Understanding Society Variable Search](https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation), and the [Documentation](https://www.understandingsociety.ac.uk/documentation).

## Using the data across multiple projects.

Adding an environment variable called `DATA_DIRECTORY` pointing at a folder, automatically saves an fst of the final dataset, so it can be loaded from anywhere using:

```{r eval = FALSE}
library(tidyusoc)
usoc <- usoc_load()
```
